---
title: "USGS Data Retrieval (API)"
author: "Clarissa Boyajian"
date: "10/6/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(janitor)
library(dataRetrieval)
library(metajam)
```

# In Class Exercise 1: USGS
```{r}
# Download the data
discharge_ventura1 <- readNWISdv(siteNumber = "11118500",
                        parameterCd = "00060",
                        startDate = "2019-10-01",
                        endDate = "2020-10-05") %>% 
  clean_names() %>% 
  rename(discharge = x_00060_00003)
```

```{r}
# Create a line plot
ggplot(data = discharge_ventura1, aes(x = date, y = discharge)) +
  geom_line()
```


#### In Class Exercise 1: Bonus
```{r}
# Download data for early years
ventura_measurement_start <- readNWISdv(siteNumber = "11118500",
                        parameterCd = "00060",
                        startDate = "1900-01-01",
                        endDate = "1980-01-01")
```

```{r}
# Find the minimum date
min(ventura_measurement_start$Date)
```

# Assignment 2.1: dataRetrieval

#### 2.1.1:
```{r}
# Download the data for multiple sites within one `readNWISdv()` function, add column with site name
discharge_2sites <- readNWISuv(siteNumber = c("11113500", "11118500"),
                        parameterCd = "00060",
                        startDate = "2021-10-03",
                        endDate = "2021-10-10",
                        tz = "PST") %>% 
  clean_names() %>% 
  rename(discharge = x_00060_00000) %>% 
  mutate(site = case_when(site_no == "11113500" ~ "Santa Paula",
                          site_no == "11118500" ~ "Ventura"))
```

#### 2.1.2:
I used the `c()` function to pull data for two different sites simultaneously. Then I cleaned up the column titles and used the `mutate()` function to added a column listing the name of each site in addition to the site number. 

#### 2.1.3:
```{r}
# Filter to only be 2021-10-04:
discharge_2021.10.04 <- discharge_2sites %>% 
  filter(date_time >= "2021-10-03 00:00:00",
         date_time <= "2021-10-06 24:59:59")

# Create a line plot of both datasets
ggplot(data = discharge_2021.10.04, 
       aes(x = date_time, 
           y = discharge, 
           colour = factor(site))) +
  geom_line() +
  scale_color_manual(values = c("Venture" = "darkgrey",
                                "Santa Paula" = "red"),
                     name = "Site") +
  labs(x = "Date",
       y = "Discharge (cubic feet per second")
```

It appears as though a stream flow peak for the Santa Paula site occurred in the early morning of October 5th. The Ventura site did not appear to have a stream flow peak in response to the storm that occurred on October 4th. 


# Assignment 2.2: Finish in class exercise 2 (DataOne)
```{r, eval=FALSE}
# Download the data packet (metadata and data files)
download_d1_data(data_url = "https://knb.ecoinformatics.org/knb/d1/mn/v2/object/urn%3Auuid%3A7fc6f6db-c5ea-426a-a743-1f2edafb43b8", path = ".")
```

```{r}
# Read the data
data_list <- read_d1_files(folder_path = "doi_10.5063_F1CJ8BPH__household_language__csv")

# Get the dataframes
hh_data <- data_list$data %>% 
  clean_names()

hh_metadata <- data_list$summary_metadata
```

```{r}
hh_data_2009to2015 <- hh_data %>% 
  filter(year >= 2009,
         year <= 2015)

percentage <- hh_data_2009to2015 %>% 
  group_by(year) %>% 
  summarize(percentage = (sum(speak_only_english) / sum(total)) * 100)
  
ggplot(percentage, aes(x = year, y = percentage)) +
  geom_point() +
  labs(x = "Year",
       y = "Percentage")
```




